---
title: "DataPreparation"
author: "Hezal Zhao"
date: "28/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
The purpose of this step is to transform the raw data into data that meets requirement for analysis, including

1.Deriving variables: some deficits measure symptoms, which may be related with other diseases. In this case, this 
type of deficits needs to be distinguished with the disease such that the measure does not overlap.


2.Recoding/rescaling: this involved reassigning values of variables. Different variables have different value scale, some variables are binary having either value “0”(No/Good; a negative result of a deficit) or “1” (Yes/Bad; a positive result of a deficit), some variables are ordinal having values 1 to 4 or 1 to 5, for convenience of analysis, all values are rescaled within 0 and 1. For example, hearing has original value “1”,”2”,”3”,”4”,”5”, it is rescaled as “0”,”0.25”,”0.5”,”0.75”,”1”. A higher value represents a worse result of a frailty deficit.


```{r}
library("dplyr")
library("tidyverse")
library(haven)
library(dineq) #Decomposition of variable inequality by population sub groups.


data1<- read_sav("Wave_4_Core_Data_Phase_2.sav")


keep_var<- c( "idauniq", "hemobwa", "hemobsi", "hemobch", "hemobcs", "hemobcl"
             ,"hemobst", "hemobre", "hemobpu", "hemobli", "hemobpi"
             ,"headldr", "headlwa", "headlba", "headlea", "headlbe", "headlwc"
             ,"headlma","headlpr", "headlsh", "headlte", "headlme", "headlmo"
             ,"headlho","hehelf","psceda", "pscedb", "pscedc", "pscedd", "pscede"
             ,"pscedf","pscedg", "pscedh","hedimbp","hediman", "hedimmi", "hedimhf"
             ,"hedimar", "hedimdi", "hedimst", "hedawbp", "hedawan","hedawmi","hedawhf"
             ,"hedawar", "hedawdi", "hedawst","hediblu", "hedibas", "hedibar", "hedibos"
             ,"hedibca", "hedibpd", "hedibps", "hedibad", "hedibde","hedbwlu", "hedbwas"
             ,"hedbwar", "hedbwos","hedbwca", "hedbwpd", "hedbwps","hedbwad", "hedbwde"
             ,"heeye", "hehear", "hefla", "hefrac", "heji", "hejiaa1", "hejiaa2", "hejiaa3"
             , "hejiaa4",   "mmpain","cfdatd", "cfdatm", "cfdaty", "cfday", "cfmem"
             ,  "cflisen", "cfani"  ,"cflisd")



data2<-data1[,keep_var]
data2<-subset(data1,select = keep_var)

data2 %>% 
  rename(
     headlph=headlte ,
     headlhg=headlho 
  )


```

```{r}
table(data2$hehelf)
#--- Recode  0 "Excellent" 1 "Poor"----#
data2$hehelf<-recode(as.numeric(data2$hehelf), `1` = 0, `2` =0.25, `3`=0.5,`4`=0.75,`5`=1)
table(data2$hehelf)


#----psced 0 "Good outcome" 1 "Bad outcome"----#
data2$psceda<-recode(as.numeric(data2$psceda), `2` = 0)
data2$pscedb<-recode(as.numeric(data2$pscedb), `2` = 0)
data2$pscedc<-recode(as.numeric(data2$pscedc), `2` = 0)
data2$pscede<-recode(as.numeric(data2$pscede), `2` = 0)
data2$pscedg<-recode(as.numeric(data2$pscedg), `2` = 0)
data2$pscedh<-recode(as.numeric(data2$pscedh), `2` = 0)
data2$pscedd<-recode(as.numeric(data2$pscedd), `1` = 0,`2`=1)
data2$pscedf<-recode(as.numeric(data2$pscedf), `1` = 0, `2`=1)



```


```{r}
#-----Need to combine newly reported heart problems with those 
#-----that are fed forward. Later  need to adjust for outcomes
#-----in earlier waves too......
data2$hedimbp[data2$hedimbp<0 & data2$hedawbp==1] <-1
data2$hedimbp[data2$hedimbp==0 & data2$hedawbp==1] <-1
table(data2$hedimbp)

data2$hediman[data2$hedimam<0 & data2$hedawan==2] <-1
data2$hediman[data2$hediman==0 & data2$hedawan==2] <-1
table(data2$hediman)

data2$hedimmi[data2$hedimmi<0 & data2$hedawmi==3] <-1
data2$hedimmi[data2$hedimmi==0 & data2$hedawmi==3] <-1
table(data2$hedimmi)

data2$hedimhf[data2$hedimhf<0 & data2$hedawhf==4] <-1
data2$hedimhf[data2$hedimhf==0 & data2$hedawhf==4] <-1
table(data2$hedimhf)

data2$hedimar[data2$hedimar<0 & data2$hedawar==6] <-1
data2$hedimar[data2$hedimar==0 & data2$hedawar==6] <-1
table(data2$hedimar)

data2$hedimdi[data2$hedimdi<0 & data2$hedawdi==7] <-1
data2$hedimdi[data2$hedimdi==0 & data2$hedawdi==7] <-1
table(data2$hedimdi)

data2$hedimst[data2$hedimst<0 & data2$hedawst==8] <-1
data2$hedimst[data2$hedimst==0 & data2$hedawst==8] <-1
table(data2$hedimst)



#-----Now recode chronic conditions---#
data2$hediblu[data2$hediblu<0 & data2$hedbwlu==1] <-1
data2$hediblu[data2$hediblu==0 & data2$hedbwlu==1] <-1
table(data2$hediblu)

data2$hedibas[data2$hedibas<0 & data2$hedbwas==2] <-1
data2$hedibas[data2$hedibas==0 & data2$hedbwas==2] <-1
table(data2$hedibas)

data2$hedibar[data2$hedibar<0 & data2$hedbwar==3] <-1
data2$hedibar[data2$hedibar==0 & data2$hedbwar==3] <-1
table(data2$hedibar)

data2$hedibos[data2$hedibos<0 & data2$hedbwos==4] <-1
data2$hedibos[data2$hedibos==0 & data2$hedbwos==4] <-1
table(data2$hedibos)

data2$hedibca[data2$hedibca<0 & data2$hedbwca==5] <-1
data2$hedibca[data2$hedibca==0 & data2$hedbwca==5] <-1
table(data2$hedibca)

data2$hedibpd[data2$hedibpd<0 & data2$hedbwpd==6] <-1
data2$hedibpd[data2$hedibpd==0 & data2$hedbwpd==6] <-1
table(data2$hedibpd)

data2$hedibps[data2$hedibps<0 & data2$hedbwps==7] <-1
data2$hedibps[data2$hedibps==0 & data2$hedbwps==7] <-1
table(data2$hedibps)

data2$hedibad[data2$hedibad<0 & data2$hedbwad==8] <-1
data2$hedibad[data2$hedibad==0 & data2$hedbwad==8] <-1
table(data2$hedibad)

data2$hedibde[data2$hedibde<0 & data2$hedbwde==9] <-1
data2$hedibde[data2$hedibde==0 & data2$hedbwde==9] <-1
table(data2$hedibde)


#----------0 "Excellent" 1 "Registered blind"---------#
data2$heeye
data2$heeye<-recode(as.numeric(data2$heeye),  `1` = 0, `2` =0.2, `3`=0.4,`4`=0.6,`6`=1)




#------0 "Excellent" 1 "Poor"----#
data2$hehear<-recode(as.numeric(data2$hehear),  `1` = 0, `2` =0.25, `3`=0.5,`4`=0.75,`5`=1)

#--- 0 "Good outcome" 1 "Bad outcome"----#
data2$hefla<-recode(as.numeric(data2$hefla),  `2` = 0)
data2$hefrac<-recode(as.numeric(data2$hefrac),  `2` = 0)
data2$heji<-recode(as.numeric(data2$heji),  `2` = 0)

#--- 0 "Good outcome" 1 "Bad outcome"----#
data2$cfdatd<-recode(as.numeric(data2$cfdatd), `1` = 0, `2`=1)
data2$cfdatm<-recode(as.numeric(data2$cfdatm), `1` = 0, `2`=1)
data2$cfdaty<-recode(as.numeric(data2$cfdaty), `1` = 0, `2`=1)
data2$cfday<-recode(as.numeric(data2$cfday), `1` = 0, `2`=1)
data2$cfmem<-recode(as.numeric(data2$cfmem), `1` = 0, `2`=1)
data2$cfdatd<-recode(as.numeric(data2$cfdatd), `1` = 0, `2`=1)
  



data2$cflisen<-as.numeric(data2$cflisen)
data2$cflisenqs<-NA
n.no<-which(data2$cflisen>0)
data2$cflisenqs[n.no] <- as.vector(ntiles.wtd(x=data2$cflisen[data2$cflisen>0], n=4 ))
table(data2$cflisenqs)
#-----label  cflisenq 1 "Worst outcome" 0 "Best outcome"----#
data2$cflisenqs [data2$cflisenqs==1]=1
data2$cflisenqs [data2$cflisenqs==2]=0.66
data2$cflisenqs [data2$cflisenqs==3]=0.33
data2$cflisenqs [data2$cflisenqs==4]=0






data2$cfani<-as.numeric(data2$cfani)
data2$cfaniq<-NA
n.no<-which(data2$cfani>0)
data2$cfaniq[n.no] <- as.vector(ntiles.wtd(x=data2$cfani[data2$cfani>0], n=5 ))
table(data2$cfaniq)
#-----label  cflisenq 1 "Worst outcome" 0 "Best outcome"----#
data2$cfaniq [data2$cfaniq==1]=1
data2$cfaniq [data2$cfaniq==2]=0.75
data2$cfaniq [data2$cfaniq==3]=0.5
data2$cfaniq [data2$cfaniq==4]=0.25
data2$cfaniq [data2$cfaniq==5]=0
table(data2$cfaniq)





table(data2$cflisd)
data2$cflisd<-as.numeric(data2$cflisd)
data2$cflisdqs<-NA
n.no<-which(data2$cflisd>0)
data2$cflisdqs[n.no] <- as.vector(ntiles.wtd(x=data2$cflisd[data2$cflisd>0], n=5 ))
table(data2$cflisdqs)
#-----label  cflisenq 1 "Worst outcome" 0 "Best outcome"----#
data2$cflisdqs [data2$cflisdqs==1]=1
data2$cflisdqs [data2$cflisdqs==2]=0.75
data2$cflisdqs [data2$cflisdqs==3]=0.5
data2$cflisdqs [data2$cflisdqs==4]=0.25
data2$cflisdqs [data2$cflisdqs==5]=0
table(data2$cflisdqs)




#--- covert varuable into numeric spss lable. 
str(data2)
data2 <- data.frame(lapply(data2, function(x) ifelse(!is.na(as.numeric(x)), as.numeric(x),  x)))
str(data2)
#----Make all negative missing values equal to NA---#
data2<- data2 %>% 
  mutate_each(funs(replace(., .<0, NA)))




#---Drop variables that are not required for frailty index
drop<- c("hedawbp", "hedawan", "hedawmi"
             ,"hedawhf", "hedawar", "hedawdi", "hedawst"
             ,"hedbwlu", "hedbwas", "hedbwar", "hedbwos"
             ,"hedbwca", "hedbwpd", "hedbwps", "hedbwad"
             ,"hedbwde", "cflisen", "cfani", "cflisd")

data3<- data2[,!(names(data2) %in% drop)]
names(data3)

names(data3)[-1] <- paste(names(data3[-1]),rep(4,ncol(data3)-1), sep = "")

drop.hej<- c("hejiaa14", "hejiaa24","hejiaa34",  "hejiaa44")
data3<- data3[,!(names(data3) %in% drop.hej)]
names(data3)

data3$mmpain4[data3$mmpain4==2] <- 0

```

```{r}
#------sort data and save wave4-------#
data4 <- data3[order(data3$idauniq),]
write.csv(data4, "wave4.csv")
save(data4,file = "tdata1.Rda")
```


