---
title: "FA08/07"
author: "Hazel Zhao"
date: "30/10/2020"
output: html_document
---

```{r include = FALSE}
knitr::opts_chunk$set(echo=FALSE)
```
Group variables by their domains. The analysis is conducted within each domain.
```{r}
library("dplyr")
library("tidyverse")
library(haven)
library(dineq)
library(haven)
library("corrplot")
library(openxlsx)
library(psych)
library(polycor)


# Group variables by their domains. The analysis is conducted within each domain.

classified_list=list()
classified_list[[1]]=c("hehelf4")
#Self-rated general health



classified_list[[2]]=c("hemobwa4", "hemobsi4", "hemobch4", "hemobcs4", "hemobcl4"
             ,"hemobst4", "hemobre4", "hemobpu4", "hemobli4", "hemobpi4")
#Mobility: 10 items concerning difficulty walking, getting up, lifting, etc.




classified_list[[3]]=c("headldr4", "headlwa4", "headlba4", "headlea4", "headlbe4", "headlwc4")
#Disabilities: Activities of Daily Living (ADLs)



classified_list[[4]]=c("headlma4","headlpr4", "headlsh4", "headlph4", "headlme4", "headlmo4","headlhg4")
#Disabilities: Instrumental Activities of Daily Living (IADLs)



classified_list[[5]]=c("hedimbp4","hediman4", "hedimmi4", "hedimhf4","hedimar4", "hedimdi4","hedimst4")
#Physician diagnosed conditions:heart problem


classified_list[[6]]=c("hediblu4", "hedibas4", "hedibar4", "hedibos4","hedibca4",
                     "hedibpd4", "hedibps4", "hedibad4", "hedibde4")
#Physician diagnosed conditions:others


classified_list[[7]]=c("heeye4", "hehear4")
#Eyesight and hearing – self-reported



classified_list[[8]]=c("hefla4", "hefrac4","heji4","mmpain4")
#bone

classified_list[[9]]=c("psceda4", "pscedb4", "pscedc4", "pscedd4", "pscede4","pscedf4","pscedg4", "pscedh4")
#mental

classified_list[[10]]=c("cfdatd4", "cfdatm4", "cfdaty4", "cfday4", "cfmem4","cflisenqs4","cflisdqs4","cfaniqs4")
#congnitive function
```

```{r}
# Basic profiling 

#separate deficits variables and demographic variables
demog=c(64:66)
candi=c(1:63,83,84,87)
load("tdata1.Rda")
wave4data=tdata1
wave4data=wave4data[wave4data$r4agey>=60,candi]  #remove people under 60 years old
total_missing=sum(complete.cases(wave4data))/nrow(wave4data)
frailty_missing=sum(complete.cases(wave4data[,1:63]))/nrow(wave4data)
#casp19_missing=sum(complete.cases(wave4data[,64:82]))/nrow(wave4data)
demo_missing=sum(complete.cases(wave4data[,demog]))/nrow(wave4data)
print(paste("completed cases for all:",total_missing)) # for all variables
print(paste("frailty items completed cases:",frailty_missing))# for 62 deficits
#print(paste("quality of life completed cases:",casp19_missing))
print(paste("demographic items completed cases:",demo_missing)) # age, gender, income, wealth

 mydata=wave4data[complete.cases(wave4data),]
 dim(mydata)
save(mydata,file ="mydata.Rda")
```
When exploring missing data, it is important to address the reason of missingness. Except for deficits of pain whilst walking, missingness of the others can be assumed missing completed at random (MCAR), because the percentage of missing value is not large, ignorance of the missingness should not affect the analysis result with bias. However, the missingness of "pain whilst walking" is special, it is related to other variables, such as the participants being unable to walk, therefore this variable is not applicable, this is been measure in other deficits, in this case, the missingness is still considered ignorable and missing at random (MAR).As a result, participants having missing data is deleted and not including in the sample, which means 78.2\% eligible participants are included in the study.

Put the annotation on variables
```{r}
load("mydata.Rda")
item_an=read.xlsx("deficits ano.xlsx")
for (i in 1:62){
   attr(mydata[,i+1],"label")=item_an[i,2]
}
```

Calculate frailty score for each person -- The frailty index

The frailty index (FI) consisted of 62 deficits, including 10 mobility difficulty deficits, 13 disability deficits, 16 physician diagnosed diseases, eyesight and hearing, self-rated general health, 4 movement deficits, 8 CES-D depression items and 8 cognitive function items.

It is a gamma distribution, which is proved using a goodness of fit test with p-value $< 0.01$. 
```{r message=FALSE, warning=FALSE}
frailty_score=apply(mydata[2:63],1,mean)
print(summary(frailty_score))
hist(frailty_score,freq=F,breaks = 100)
library(ggplot2)
library(MASS)
library(goftest)



den=density(frailty_score)
dat <- data.frame(x = den$x, y = den$y)
fit.params= fitdistr(frailty_score,"gamma",lower=c(0,0)) #fit gamma distribution
ggplot(data = dat, aes(x = x,y = y)) + 
  geom_point(size = 3) +     
  geom_line(aes(x=dat$x, y=dgamma(dat$x,fit.params$estimate["shape"], fit.params$estimate["rate"])), color="red", size = 1) + 
  theme_classic()
ggplot(data = dat) +
  geom_histogram(data = as.data.frame(frailty_score), aes(x=frailty_score, y=..density..),bins = 50,fill="blue",color="black") +
  geom_line(aes(x=dat$x, y=dgamma(dat$x,fit.params$estimate["shape"], fit.params$estimate["rate"])), color="red", size = 1) + 
  theme_classic()+
  labs(x="Frailty Index",y="Density")

summary(frailty_score)
sd(frailty_score)

# The scale and shape of "gamma" districution
scale=sd(frailty_score)^2/mean(frailty_score)
shape=mean(frailty_score)^2/sd(frailty_score)^2
```



