---
title: "FA08/07"
author: "Hazel Zhao"
date: "30/10/2020"
output: html_document
---

```{r include = TRUE}
knitr::opts_chunk$set(echo=TRUE)
```

```{r}
library("dplyr")
library("tidyverse")
library(haven)
library(dineq)
library(haven)
library("corrplot")
library(openxlsx)
library(psych)
library(polycor)
```

3. Diagonosed condition: CVD+chronic

Diagnosed conditions include 7 cardiovascular disease (CVD), 9 chronic diseases.

```{r}
var_list=unlist(classified_list[5:6])
temp=data.frame(mydata[,var_list])
diagnoed=matrix(nrow=length(var_list),ncol=6,dimnames = list(var_list,c("Mean","SD","Min","Max","percen of'0'","percen of'1'")))


  for (i in 1: length(var_list)){
    tab=table(temp[,i])
   diagnoed[i,5]=tab[[1]]/sum(tab)
    diagnoed[i,6]=tab[[2]]/sum(tab)
    diagnoed[i,1]=mean(temp[,i])
    diagnoed[i,2]=sd(temp[,i])
   diagnoed[i,3]=min(temp[,i])
    diagnoed[i,4]=max(temp[,i])
  }

diagnoed=data.frame(diagnoed)
diagnoed
for (i in 1:ncol(temp)){
  print(attr(temp[,i],"label"))
}

diagnoed_N=apply(mydata[,var_list],1,sum)
ggplot(NULL,aes(x=as.factor(diagnoed_N)))+ 
  geom_bar(aes(y=(..count..)/sum(..count..)))+
  labs(x="number of disease",y="percentage of participants",title = "Diagnosed Condition")+
  theme(plot.title = element_text(hjust = 0.5),panel.grid =element_blank()) 
```
Since CVD and chronic diseases are different type of diseases, the correlation of the items in the two groups are also low, factor analysis is considered to be conducted within the two groups separately.

The factor adequacy index of chronic diseases is lower than 0.5, so EFA is not plausible.

As for the number of factors for CVD items, Kaiser's rule and and scree plot (2 factors) is contradict with parallel analysis (4 factors). Similar procedure is conducted as previously, performance of models with factor numbers of 2 to 4 are compared, the two factors structure is optimal.

```{r}
#drop=c("hedimar4")
#temp=temp[,!(names(temp) %in% drop)]
#corre_diagnoed=cor(temp)
corre_diagnoed=tetrachoric(temp[,1:7],smooth=T)
corre_diagnoed=pmax(corre_diagnoed$rho,t(corre_diagnoed$rho))
KMO(corre_diagnoed)
bart_spher(corre_diagnoed)
```
```{r}
library("corrplot")
corrplot(corre_diagnoed, is.corr = TRUE, number.cex = .5, tl.col = 'black',order = 'hclust', addrect = 4, tl.cex = 0.7)
corrplot.mixed(corre_diagnoed, is.corr = TRUE, number.cex = .5, tl.col = 'black',tl.cex = 0.4)
```
```{r}
eigen.corre <- eigen(corre_diagnoed)
eigen.corre$values
```
```{r}
scree(corre_diagnoed)
VSS.scree(corre_diagnoed)
```
```{r}
fa.parallel(corre_diagnoed, n.obs= nrow(mydata),fm = "pa",fa="both",n.iter = 1000)
#VSS(corre_diagnoed,n=3,plot = T,n.obs= nrow(mydata))
```

```{r}

diagnoed.fa3<- fa(corre_diagnoed, nfactors = 2,n.obs=4800,rotate = 'oblimin',scores = 'regression', fm = 'pa')
print(diagnoed.fa3,sort = T)
fa.diagram(diagnoed.fa3, cut=0.299,labels=names(diagnoed.fa),digits=2,e.size=0.05,rsize=0.1,cex=1.2,adj=1,marg=c(1,1,1.5,1))
cor.plot(diagnoed.fa3$Phi,main = "Factor Correlations",labels = c("Heart disease","Other CVD"),cex = 1.5)

f1=c(2,3,4,5)
f2=c(1,6,7)
Heart_diseases=as.matrix(temp[,f1])%*%diagnoed.fa3$loadings[1:4,1]
Other_CVD=as.matrix(temp[,f2])%*%diagnoed.fa3$loadings[5:7,2]
score3=factor.scores(temp[,1:7],diagnoed.fa3)
cvd_factor=score3$scores
```
The two factors are heart diseases including items: Angina heart attack, Congestive heart failure and Abnormal heart rhythm. Diabetes or high blood sugar, stroke and high blood pressure are factorised as other CVD.

The total observed variance is 47\%, which is relatively low compared with previous models for other items. Consistently, TLI is lower than the cutoff value 0.8 and RMSEA is also high, p-value of likelihood chi square suggests that the model is sufficient, though it dose not fit the data very well. Sole reliance on one or two fit index seems imprudent, the chi squared goodness of fit and the two factors is meaningful, so the two factors solution is acceptable. 


4. self-reported items (only EDA)
```{r}
var_list=c(unlist(classified_list[1]),unlist(classified_list[7]))
temp=data.frame(mydata[,var_list])
self_reported_item=matrix(nrow=length(var_list),ncol=7,dimnames = list(var_list,c("Mean","SD","%oflow ","%of2nd","%of3rd","%of4th","%ofhigh")))

  for (i in 1: length(var_list)){
    tab=table(temp[,i])
   self_reported_item[i,3]=tab[[1]]/sum(tab)
   self_reported_item[i,4]=tab[[2]]/sum(tab)
   self_reported_item[i,1]=mean(temp[,i])
   self_reported_item[i,2]=sd(temp[,i])
   self_reported_item[i,5]=tab[[3]]/sum(tab)
   self_reported_item[i,6]=tab[[4]]/sum(tab)
   self_reported_item[i,7]=tab[[5]]/sum(tab)
  }
self_reported_item=data.frame(self_reported_item)
self_reported_item
for (i in 1:ncol(temp)){
  print(attr(temp[,i],"label"))
}

ggplot(NULL,aes(x=as.factor(temp[,1])))+ 
  geom_bar(aes(y=(..count..)/sum(..count..)))+
  labs(x="rank",y="percentage of participants",title = "self-reported general health")+
  theme(plot.title = element_text(hjust = 0.5),panel.grid =element_blank()) 
ggplot(NULL,aes(x=as.factor(temp[,2])))+ 
  geom_bar(aes(y=(..count..)/sum(..count..)))+
  labs(x="rank",y="percentage of participants",title = "self-reported eyesight")+
  theme(plot.title = element_text(hjust = 0.5),panel.grid =element_blank()) 
ggplot(NULL,aes(x=as.factor(temp[,3])))+ 
  geom_bar(aes(y=(..count..)/sum(..count..)))+
  labs(x="rank",y="percentage of participants",title = "self-reported hearing")+
  theme(plot.title = element_text(hjust = 0.5),panel.grid =element_blank()) 

```
