---
title: "FA08/07"
author: "Hazel Zhao"
date: "30/10/2020"
output: html_document
---

```{r include = TRUE}
knitr::opts_chunk$set(echo=TRUE)
```

```{r}
library("dplyr")
library("tidyverse")
library(haven)
library(dineq)
library(haven)
library("corrplot")
library(openxlsx)
library(psych)
library(polycor)
```


5. Movement items
```{r}
var_list=c(unlist(classified_list[8]))
temp=data.frame(mydata[,var_list])
fall_=matrix(nrow=length(var_list),ncol=6,dimnames = list(var_list,c("Mean","SD","Min","Max","percen of'No'","percen of'Yes'")))


  for (i in 1: length(var_list)){
    tab=table(temp[,i])
    fall_[i,5]=tab[[1]]/sum(tab)
    fall_[i,6]=tab[[2]]/sum(tab)
    fall_[i,1]=mean(temp[,i])
    fall_[i,2]=sd(temp[,i])
    fall_[i,3]=min(temp[,i])
    fall_[i,4]=max(temp[,i])
  }

fall_=data.frame(fall_)
fall_
for (i in 1:ncol(temp)){
  print(attr(temp[,i],"label"))
}

```
```{r}
corre_diagnoed=tetrachoric(temp)
corre_diagnoed=pmax(corre_diagnoed$rho,t(corre_diagnoed$rho))
KMO(corre_diagnoed)
bart_spher(corre_diagnoed)
pc=fa(corre_diagnoed, nfactors = 1,n.obs=5500,rotate = 'none',scores = 'regression', fm = 'pa')
print(pc,sort=T)
f1=c(2,3,1)
Movement=as.matrix(temp[,f1])%*%pc$loadings[1:3,1]
score4=factor.scores(temp,pc)
Movement_factor=score4$scores

```
Both factor adequacy methods suggest that the correlation of the 4 movement items is sufficient to conduct factor analysis, though 0.58 may be a low value of factor adequacy index. There are only 4 items, a 1 factor is supposed to extracted based on Kaiser's rule, scree plot and interpretability. Item pain whilst walking is cut off in the model because of the low factor loading.



6. mental health: CESD
```{r}
var_list=unlist(classified_list[9])
temp=data.frame(mydata[,var_list])
CESD=matrix(nrow=length(var_list),ncol=6,dimnames = list(var_list,c("Mean","SD","Min","Max","percent of'0'","percent of'1'")))


  for (i in 1: length(var_list)){
    tab=table(temp[,i])
   CESD[i,5]=tab[[1]]/sum(tab)
   CESD[i,6]=tab[[2]]/sum(tab)
   CESD[i,1]=mean(temp[,i])
   CESD[i,2]=sd(temp[,i])
   CESD[i,3]=min(temp[,i])
   CESD[i,4]=max(temp[,i])
  }

CESD=data.frame(CESD)
CESD
for (i in 1:ncol(temp)){
  print(attr(temp[,i],"label"))
}
mean_CESD=apply(mydata[,var_list],1,mean)
  ggplot(NULL,aes(x=mean_CESD))+ 
  geom_bar(aes(y=(..count..)/sum(..count..)))+
  labs(x="CESD",y="Number of observations",title = "CESD score")+
  theme(plot.title = element_text(hjust = 0.5),panel.grid =element_blank())  
```
 
 
```{r}
corre_mental=tetrachoric(mydata[,var_list],smooth=T)
corre_mental=pmax(corre_mental$rho,t(corre_mental$rho))
KMO(corre_mental)
bart_spher(corre_mental)
corrplot(corre_mental, is.corr = TRUE, number.cex = .5, tl.col = 'black',order = 'hclust', addrect = 2, tl.cex = 0.7)
corrplot.mixed(corre_mental, is.corr = TRUE, number.cex = .5, tl.col = 'black',tl.cex = 0.6)
```
```{r}
eigen.corre <- eigen(corre_mental)
eigen.corre$values
```
```{r}
scree(corre_mental)
```
```{r}
fa.parallel(corre_mental,fm="pa",fa="fa",n.obs= nrow(mydata),n.iter = 1000)
VSS(corre_mental,n.obs = 5500,n=3,plot = T)
```
Although there is only one eigenvalues of the correlation coefficient matrix larger than 1, and 1 point above the elbow of the scree plot, parallel analysis have 2 eigenvalues of the actual data above the random data, suggesting a 2 factors model.

```{r}

mental.fa3<- fa(corre_mental, nfactors = 2, rotate = 'oblimin',n.obs=5500,scores = 'regression', fm = 'pa')
print(mental.fa3,sort = T)
factor.plot(mental.fa3, labels = rownames(mental.fa3$loadings))
fa.diagram(mental.fa3, cut=0.299,labels=names(mental.fa3),digits=2,e.size=0.05,rsize=0.1,cex=1.2,adj=1,marg=c(1,1,1.5,1))
cor.plot(mental.fa3$Phi,main = "Factor Correlations",labels = c("Factor 1","Factor 2"),cex = 1.5)
f1=c(4,6,7,5,1)
Absence_positive_anhodonia=as.matrix(temp[,f1])%*%mental.fa3$loadings[1:5,1]
Somatic_activity=as.matrix(temp[,-f1])%*%mental.fa3$loadings[6:8,2]
score5=factor.scores(temp,mental.fa3)
mental_factor=score5$scores
```
CES-D depression scales have two constructs: absence of positive affect or anhedonia, somatic activity or inactivity. Absence of positive affect or anhedonia is measured by "happy" ,"enjoyed life","felt sad","felt lonely","felt depressed". These items are  distinct from the other items (felt everything was an effort, could not get going, felt restless) measuring the other factor of CES-D. The underlying structure is clear. It is noticeable that the two factors are significantly related though they are conceptually distinctly defined. The interaction between positive or negative affect (PA/NA) and somatic activity (SA) has been studied by researchers.

```{r}
f1=c(1,4:7)
score=apply(temp[,f1],1,sum)
hist(score,freq=F)
score=as.matrix(temp[,f1])%*%mental.fa3$loadings[f1,1]
hist(score,freq=F)
score=apply(temp[,-f1],1,sum)
hist(score,freq=F)
score=as.matrix(temp[,-f1])%*%mental.fa3$loadings[-f1,2]
hist(score,freq=F)
```


