---
title: "FA08/07"
author: "Hazel Zhao"
date: "30/10/2020"
output: html_document
---

```{r include = TRUE}
knitr::opts_chunk$set(echo=TRUE)
```

```{r}
library("dplyr")
library("tidyverse")
library(haven)
library(dineq)
library(haven)
library("corrplot")
library(openxlsx)
library(psych)
library(polycor)
```

Factor analysis is conducted for the items within domains. For all domains, Principal axis method is used to extract factors. For model with more than one factor, factor rotation method is oblique to generate simpler structure.

Note: generally, if we need conduct analysis repeatedly, making code reusable (define functions) is better practice. However, this analysis is pretty exploratory(e.g.,dropping items based on the result of the analysis). Being flexible is optimal.

1. Mobility items
```{r}
var_list=unlist(classified_list[2])
temp=data.frame(mydata[,var_list])
mobility_difficulty=matrix(nrow=length(var_list),ncol=6,dimnames = list(var_list,c("Mean","SD","Min","Max","percent of'0'","percent of'1'")))

for (j in 1:4){
  for (i in 1: length(var_list)){
    tab=table(temp[,i])
    mobility_difficulty[i,5]=tab[[1]]/sum(tab)
    mobility_difficulty[i,6]=tab[[2]]/sum(tab)
    mobility_difficulty[i,1]=mean(temp[,i])
    mobility_difficulty[i,2]=sd(temp[,i])
    mobility_difficulty[i,3]=min(temp[,i])
    mobility_difficulty[i,4]=max(temp[,i])
  }
}
mobility_difficulty=data.frame(mobility_difficulty)
mobility_difficulty
for (i in 1:ncol(temp)){
  print(attr(temp[,i],"label"))
}
mobility_diff=apply(mydata[,var_list],1,sum)/10
ggplot(NULL,aes(x=as.factor(mobility_diff)))+ 
  geom_bar(aes(y=(..count..)/sum(..count..)))+
  labs(x="number of mobility difficulties",y="percentage of participants",title = "Mobility difficulty")+
  theme(plot.title = element_text(hjust = 0.5),panel.grid =element_blank())
```
```{r}
# tetrachoric coefficient for binary data


library(REdaS)
corre_mobility=tetrachoric(temp,smooth=T)
corre_mobility=pmax(corre_mobility$rho,t(corre_mobility$rho))

print("coefficient of Mobility Items")
corre_mobility

print ("factor adequacy test")
KMO(corre_mobility)
bart_spher(corre_mobility)
```
The adequacy index of Kaiser-Meryer-Olkin (KMO) method is larger than 0.5. Additionally, Bartlett's Test of Sphericity is significant (the chi square's p-value $< 0.01$). Both tests indicate that the relationship between variables is strong enough, the data is sufficient to conduct factor analysis.


```{r}
# draw coefficient
library("corrplot")
corrplot(corre_mobility, is.corr = TRUE, number.cex = .5, tl.col = 'black',order = 'hclust', addrect = 2, tl.cex = 0.7)
corrplot.mixed(corre_mobility, is.corr = TRUE, number.cex = .7, tl.col = 'black',tl.cex = 0.5)
```
Determining the number of factors:

There is one eigenvalue larger than 1, which means according to Kaiser's rule, 1 factor should be extracted. This is consistent with the scree plot. However, parallel analysis, the most commonly used method, shows 3 factors can be retained.


```{r}
#eigenvalue
eigen.corre <- eigen(corre_mobility)
eigen.corre$values
#scree plot
scree(corre_mobility,factors = T,pc=F)
#parallel analysis
fa.parallel(corre_mobility,n.obs= 5500,fm="pa",fa="both",n.iter = 1000)
```
3 factors, difficulty related to upper limb, difficulty related to lower limb, other difficulty, are retained for the 10 Mobility items. Because all the items measures the same domain and strongly correlated, 1 factor model may also make sense. However one factor model does not add extra information for discovering the underlying structure of this domain since the observed variables have been categorized as mobility. Meanwhile parallel analysis and interpretability also support a 3 factors model and therefore the 3 factors model is preferred. These items are highly loaded on their own underlying factor, showing a good measure for the sub-aspect of mobility.

```{r}
# fit model
mobility.fa<- fa(corre_mobility, nfactors = 3, n.obs=5520,rotate = 'oblimin',scores = T, fm = 'pa')
print(mobility.fa,sort = T)

#draw diagram of the factor structure and correlation between factors
fa.diagram(mobility.fa, cut=0.3,labels=names(mobility.fa),digits=2,e.size=0.05,rsize=0.1,cex=1.2,adj=1,marg=c(.5,.5,1,.5))
cor.plot(mobility.fa$Phi,main = "Factor Correlations",labels = c("Upper limb","Other","Lower limb"),cex = 1.3)

```


```{r}

# calculate factor score 
f1=c(8,9,7)
f2=c(4,5,1)
f3=c(3,2,6,10)
Upper_limb=as.matrix(temp[,f1])%*%mobility.fa$loadings[1:3,1]
Lower_limb=as.matrix(temp[,f2])%*%mobility.fa$loadings[4:6,2]
Other=as.matrix(temp[,f3])%*%mobility.fa$loadings[7:10,3]
score1=factor.scores(temp,mobility.fa)
mobility_factor=score1$scores
```

2. ADLs+IADLs
```{r}
var_list=unlist(classified_list[3:4])
var_list
temp=data.frame(mydata[,var_list])
Disability=matrix(nrow=length(var_list),ncol=6,dimnames = list(var_list,c("Mean","SD","Min","Max","percent of'0'","percent of'1'")))

for (j in 1:4){
  for (i in 1: length(var_list)){
    tab=table(temp[,i])
    Disability[i,5]=tab[[1]]/sum(tab)
    Disability[i,6]=tab[[2]]/sum(tab)
    Disability[i,1]=mean(temp[,i])
    Disability[i,2]=sd(temp[,i])
    Disability[i,3]=min(temp[,i])
    Disability[i,4]=max(temp[,i])
  }
}
Disability=data.frame(Disability)
Disability
for (i in 1:ncol(temp)){
  print(attr(temp[,i],"label"))
}
disability=apply(mydata[,var_list],1,sum)
ggplot(NULL,aes(x=as.factor(disability)))+ 
  geom_bar(aes(y=(..count..)/sum(..count..)))+
  labs(x="number of disability",y="percentage of participants",title = "Disability")+
  theme(plot.title = element_text(hjust = 0.5),panel.grid =element_blank()) 
```
```{r}
#coefficient
#drop=c("headlpr4","headlsh4")
drop=c("headlpr4") #drop prepare hot meal since its cross loading on both factors
temp=temp[,!(names(temp) %in% drop)]
corre_disability=tetrachoric(temp,smooth=T)
corre_disability=pmax(corre_disability$rho,t(corre_disability$rho))
#corre_disability=cor(temp)

#factor adequacy
KMO(corre_disability)
bart_spher(corre_mobility)
```
```{r}
library("corrplot")
corrplot(corre_disability, is.corr = TRUE, number.cex = .5, tl.col = 'black',order = 'hclust', addrect = 2 , tl.cex = 0.7)
corrplot.mixed(corre_disability, is.corr = TRUE, number.cex = .5, tl.col = 'black', tl.cex = 0.5)
```

# decide number of factors
```{r}
eigen.corre <- eigen(corre_disability)
eigen.corre$values
```
```{r}
scree(corre_disability)
```
```{r}
fa.parallel(corre_disability, n.obs= nrow(mydata),fm="pa",fa="both",n.iter = 1000)

```
```{r}
disability.fa1<- fa(corre_disability, nfactors = 2, n.obs=5500,rotate = 'none',scores = 'regression', fm = 'pa',oblique.scores=TRUE)
#disability.fa1
print(disability.fa1,sort = T)
disability.fa3<- fa(corre_disability, nfactors = 2,n.obs=5500,rotate = 'oblimin',scores = 'regression', fm = 'pa',oblique.scores=TRUE )
#disability.fa3
print(disability.fa3,sort = T)
factor.plot(disability.fa3, labels = rownames(disability.fa3$loadings))
fa.diagram(disability.fa3, cut=0.299,labels=names(disability.fa3),digits=2,e.size=0.05,rsize=0.1,cex=1.2,adj=1,marg=c(1,1,1.5,1))
cor.plot(disability.fa3$Phi,main = "Factor Correlations",labels = c("ADLs disability","IADLs disability"),cex = 1.3)

```
The factor structure of disability items is clear because it uses well defined ADLs and IADLs to measure disability and this structure also been proved by the statistical approach, methods for determining number of factors and model fitting. ADLs measures more difficulty with physical activities, such as getting in and out of bed, using the toilet, dressing, bathing or showering, eating. IADLs measures more complex activities and are more related to cognitive function than physical function, which is distinct from ADLs.

```{r}
f1=c(5,6,12,1,2,3,8,4)
Physical_disability=as.matrix(temp[,f1])%*%disability.fa3$loadings[1:8,1]
Cognitive_disability=as.matrix(temp[,-f1])%*%disability.fa3$loadings[9:12,2]
score2=factor.scores(temp,disability.fa3)
disability_factor=score2$scores
ADLs_disability=Physical_disability
IADLs_disability=Cognitive_disability
```
